A Monte Carlo Synthetic-Acceleration Method for Solving the Thermal Radiation Diffusion Equation\tnoteref{ornl-cr}

### 2013-02-19 doc/jcomp2012/jcomp2012.tex
adding plots for sequential vs mcsa convergence.

\subsection{Comparison to Sequential Monte Carlo}
\label{subsec:sequential comparison}
To further motivate using Monte Carlo Synthetic Acceleration, we
compare its performance to Halton's Sequential Monte Carlo method on
which our previous work in this area was based. For this comparison,
we use the same transient Poisson problem as described in the previous
section and choose only the 5-point stencil to discretize the
Laplacian operator as the previous results yielded little qualitative
difference between the discretizations. Both MCSA and Halton's method
are used with the adjoint Monte Carlo solver. In order to complete the
same study as in the previous section, the number of histories
computed by the Monte Carlo solver at each iteration had to be doubled
to $2 \times N \times N$ in order to ensure convergence and a
monotonically decreasing residual in Sequential Monte Carlo
Method. Figure~\ref{fig:seq_cpu_time} gives the CPU time results for
this comparison as a function of problem size while
Figure~\ref{fig:seq_iterations} gives the number of iterations to
converge as a function of problem size. In both cases, using the Monte
Carlo solver as a synthetic acceleration rather than in a pure
residual Monte Carlo scheme resulted in a reduction in both CPU time
and iterations required to converge. The additional Richardson
extrapolation between each Monte Carlo solve in the MCSA method gives
a better converged residual source to use with the Monte Carlo
calculation while the Sequential method requires more iterations to
achieve the same level of convergence in the residual.

@@

The benefits of using a synthetic acceleration scheme are also noted
when the infinity norm of the residual computed computed at each
iteration for both methods was collected at each iteration for a fixed
problem sizes of $N=100$ and $N=500$ as shown in figures
Figure~\ref{fig:seq_100} and \ref{fig:seq_500} respectively. In both
cases, the Sequential method is subject to two regimes of exponential
convergence with the later regime converging the slowest while the
MCSA method exhibits a single rate of exponential convergence, much
higher than that computed by Halton's method.

>>>>

\subsection{Comparison to Sequential Monte Carlo}
\label{subsec:sequential comparison}
To further motivate using Monte Carlo Synthetic Acceleration, we
compare its performance to Halton's Sequential Monte Carlo method on
which our previous work in this area was based. For this comparison,
we use the same transient Poisson problem as described in the previous
section and choose only the 5-point stencil to discretize the
Laplacian operator as the previous results yielded little qualitative
difference between the discretizations. Both MCSA and Halton's method
are used with the adjoint Monte Carlo solver. In order to complete the
same study as in the previous section, the number of histories
computed by the Monte Carlo solver at each iteration had to be doubled
to $2 \times N \times N$ in order to ensure convergence in Sequential
Monte Carlo Method. Figure~\ref{fig:seq_cpu_time} gives the CPU time
results for this comparison as a function of problem size while
Figure~\ref{fig:seq_iterations} gives the number of iterations to
converge as a function of problem size. In both cases, using the Monte
Carlo solver as a synthetic acceleration rather than in a pure
residual Monte Carlo scheme resulted in a reduction in both CPU time
and iterations required to converge. The additional Richardson
extrapolation between each Monte Carlo solve in the MCSA method gives
a better converged residual source to use with the Monte Carlo
calculation while the Sequential method requires more iterations to
achieve the same level of convergence in the residual.

@@

The benefits of using a synthetic acceleration scheme are also noted
when the infinity norm of the residual computed computed at each
iteration for both methods was collected at each iteration for a fixed
problem sizes of $N=100$ and $N=500$ as shown in figures
Figure~\ref{fig:seq_100} and \ref{fig:seq_500} respectively. In both
cases, the Sequential method is subject to two regimes of exponential
convergence with the later regime converging the slowest while the
MCSA method exhibits a single rate of exponential convergence, much
higher than that computed by Halton's method. Even with the doubling
of the number of stochastic histories computed per time step in order
to ensure convergence for the Sequential method, we still see
robustness issues with a non-monotonically decreasing residual
observed in both cases and more pronounced for the $N=100$ case. In
both cases the MCSA solver is observed to be robust with a
monotonically decreasing residual.
